Categorical Predictor Embeddings in Actuarial Modeling

### 2021-03-22 manuscript/manuscript.Rmd
update pdf

---
title: Categorical Predictor Embeddings in Actuarial Modeling
authors:
  - name: Kevin Kuo 
    affiliation: RStudio and Kasa AI
    email: kevin@kasa.ai
  - name: Ronald Richman
    affiliation: QED Actuaries and Consultants
    email: ronaldrichman@gmail.com
abstract: |
  Enter the text of your abstract here.
keywords:
  - blah
  - blee
  - bloo
  - these are optional and can be removed
bibliography: references.bib
output:
  bookdown::pdf_book:
    base_format: rticles::arxiv_article
header-includes: |
    \usepackage{amsmath}
    \usepackage{mathtools}
    \usepackage{bm}
    \usepackage{bbm}
    \usepackage{float}
    \DeclarePairedDelimiter\ceil{\lceil}{\rceil}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.pos = "!h", out.extra = "")
```

@@

Recently, several studies of insurance problems have applied an alternative approach, originally from the natural 
known as **categorical embeddings**. Instead of trying to capture the differences between categories using a single 
factor, the categorical embedding approach rather maps each category to a low-dimensional numeric vector, which is then
used within the model as a new predictor variable.

@@

@@

```{=latex}
\begin{equation}
  z^1 = \sigma(a_1.X'+ b_1), 
\end{equation}
```
where we represent the output of embedding layers concatenated together with any numerical inputs as $X'$.

@@

Attention mechanisms have been widely applied within the deep learning literature to give more flexibility to 
RNNs [@Bahdanau2015a] and subsequently as a replacement for FCNs in the so-called Transformer architecture, which is 
now widely used in natural language processing [@Vaswani2017a], and, more recently, has been applied to computer vision 
and tabular modelling tasks. Attention mechanisms allow deep neural nets to choose covariates entering the model in a 
flexible manner; these types of applications are demonstrated in, for example, [@Vaswani2017a]. In this section we 
provide an introduction to attention mechanisms and then discuss applications of attention to modeling tabular data.

@@

```{=latex}
\begin{equation}
A: \mathbb{R}^{q \times (q \times d) \times (q \times d)} \to \mathbb{R}^q, ~~
(\mathbf{q}, K, V) ~\mapsto ~ \mathbf{v} = A(\mathbf{q}, K, V),
\label{eq:attention}
\end{equation}
```

where $\mathbf{q} \in \mathbb{R}^q$ is a query vector, $K \in \mathbb{R}^{q \times d}$ is a matrix of key vectors,
$V \in \mathbb{R}^{q \times d}$ is a matrix of value vectors and $\mathbf{v} \in \mathbb{R}^q$, the output of the 
attention mechanism, is a new vector of values. Several different options for the mapping \eqref{eq:attention} exist in 
the literature; here we first describe the self-attention mechanism used in [@Vaswani2017a]. In more detail, consider 
that $V$ is a matrix of covariates relating to a regression problem, which can be composed of embeddings of the same 
dimension for several categorical variables. We wish to apply the attention mechanism to $V$ to estimate the most 
important of these embeddings for the regression problem. To do this, we formulate the query vector $\mathbf{q}$ which 
contains information about the regression problem at hand; this is usually done by passing $L$ through a FCN. 
Similarly, the matrix $K$ contains the relevance of each row of $L$ for the regression problem and is also usually 
derived by passing each row of $L$ through another FCN. Finally, the matrix $V$ can be set equal to $L$ or 
alternatively, can be derived as a transformation of $L$ using a third FCN. The first step of the attention mechanism 
is to derive a vector $\mathbf{a}$ of attention scores:

```{=latex}
\begin{equation}
\mathbb{R}^{(1 \times q) \times (q \times d)} \to \mathbb{R}^{1 \times d}, ~~
(\bold{q}, K) ~\mapsto ~ \mathbf{a} = \langle \mathbf{q}, K^T \mathbb\rangle.
\label{eq:attention_scores}
\end{equation}
```
The attention scores are then normalized using a soft max function i.e.

```{=latex}
\begin{equation}
a^*_i = \frac{e^{(a_i)}}{\sum_{n=1}^{d} e^{(a_n)}}.
\end{equation}
```
Finally, a linear combination of $V$ is formed:

```{=latex}
\begin{equation}
\mathbb{R}^{(1 \times d) \times (d \times q)} \to \mathbb{R}^{1 \times q}, ~~
(\mathbf{a^*}, V) ~\mapsto ~ \mathbf{v} = \langle \mathbf{a^*}, V \mathbb\rangle.
\label{eq:attention_output}
\end{equation}
```

@@

These model architectures are relatively uncomplicated by today's standards, and are so chosen to better highlight the
embedding components. Later, in Section \ref{attn}, we will investigate more involved architectures utilizing embeddings
that represent the state of the art for modeling tabular data.

@@

-   Target variable: Amount paid on building claim
-   Predictors:
    - $\log$ `building_insurance_coverage`
    - `basement_enclosure_type`, 
    - `number_of_floors_in_the_insured_building`
    - *prefix* of `flood_zone`
    - `primary_residence`
-   Link function: $\log$
-   Distribution: gamma

We take the log of the continuous predictor `building_insurance_coverage`, which allows the scale of the predictor to 
match that of the target variable. Because the `flood_zone` variable in the original data contains 60 levels, we take 
the prefix of the zone code, which corresponds to the level of risk as determined by FEMA [^8]. For example, `"A01"`, 
`"A02"`, and so on are recoded as simply `"A"`. A log link together with the gamma distribution is a standard choice 
for severity modeling, which provides a multiplicative structure where the response is positive. Following the notation
in \@ref(catdata), the number of parameters in the GLM is

\begin{equation}
1 + |J^{\text{num}}| + \left(\sum_{j \in J^{\text{cat}}} n_{\mathcal{P}}^j\right) - |J^{\text{cat}}|:
(\#eq:glm1)
\end{equation}

one for the intercept, one for each numeric variable, and the numbers of levels minus one (due to dummy coding) for 
each categorical variable.

[^8]: https://www.floodsmart.gov/flood-map-zone/find-yours

@@

Table: (\#tab:relativities) Excerpt of relativities for Model 1.

|variable       |label                                       | relativity|
|:--------------|:-------------------------------------------|----------:|
|occupancy_type |Single family residence                     |       1.00|
|occupancy_type |Non-residential building                    |       1.37|
|occupancy_type |2 to 4 unit residential buidling            |       1.15|
|occupancy_type |Residential building with more than 4 units |       1.64|
|flood_zone     |B                                           |       1.00|
|flood_zone     |A                                           |       0.84|
|flood_zone     |C                                           |       0.60|
|flood_zone     |X                                           |       0.70|
|flood_zone     |V                                           |       0.78|
|flood_zone     |D                                           |       0.89|

@@

For Model 2, we build a feedforward neural network, whose architecture is shown in Figure \@ref(fig:nn-1). 
The categorical inputs go through one-dimensional embedding
layers where they are each mapped to a scalar. The embeddings are then concatenated with the numeric predictors, which
have been normalized in data pre-processing, before being sent to a hidden linear layer of 64 units. A ReLU activation,

\begin{equation}
x \mapsto \max(0, x),
\end{equation}

is then applied to the outputs before they are sent to a final output layer resulting in a scalar. To constrain
predictions to be positive, we apply softplus activation, which is defined to be

\begin{equation}
x \mapsto \log(1 + e^x). 
\end{equation}

```{r nn-1, echo = FALSE, fig.cap = "Architecture of Models 2 and 4.", out.width = "80%", fig.align = "center"}
knitr::include_graphics("images/nn.png")
```

@@

In Tables \@ref(tab:embeddings-occupancy-type) and \@ref(tab:embeddings-flood-zone), we show the learned embeddings
for the `occupancy_type` and an excerpt of them for `flood_zone`. While it may be tempting to compare these values
to those in Table \@ref(tab:relativities), they represent fundamentally different concepts. It is helpful to think
of the embedding values for each variable as a continuous represention of the variable itself, rather than a 
"coefficient" associated with the variable. Note that it is not straightforward to inspect the embeddings from a
table and infer relationships to the response variable, even in a shallow network such as ours, due to nonlinearities
within the model. Also note that we now have a way to incorporate the full granularity of the `flood_zone` variable,
which is why more levels are shown in Table \@ref(tab:embeddings-flood-zone) as compared to Table 
\@ref(tab:relativities). In Section \ref{interpretation}, we will discuss techniques to visualize these embeddings.

Table: (\#tab:embeddings-occupancy-type) Learned embeddings for `occupancy_type`.

|level                                       | embedding|
|:-------------------------------------------|---------:|
|2 to 4 unit residential buidling            |      6.69|
|Non-residential building                    |      7.16|
|Residential building with more than 4 units |      6.33|
|Single family residence                     |      5.43|

Table: (\#tab:embeddings-flood-zone) Excerpt of learned embeddings for `flood_zone`.

|level | embedding|level | embedding|level | embedding|level | embedding|level | embedding|
|:-----|---------:|:-----|---------:|:-----|---------:|:-----|---------:|:-----|---------:|
|A     |      0.38|A09   |     -1.71|A18   |     -1.40|A99   |     -2.68|V     |     -0.41|
|A00   |     -0.35|A0B   |      3.85|A19   |     -1.70|AE    |     -4.46|V02   |      0.65|
|A01   |     -6.76|A10   |     -4.03|A20   |     -4.05|AH    |     -2.35|V03   |     -0.29|
|A02   |     -4.28|A11   |     -5.32|A21   |     -2.08|AHB   |     -1.71|V04   |     -1.15|
|A03   |     -5.93|A12   |     -3.23|A22   |     -2.11|AO    |     -6.07|V05   |     -0.50|
|A04   |     -7.82|A13   |      2.03|A23   |     -1.40|AOB   |     -4.56|V06   |     -1.92|
|A05   |     -3.42|A14   |     -5.72|A24   |      2.23|AR    |     -0.48|V07   |     -2.91|
|A06   |     -4.97|A15   |     -3.67|A25   |      0.28|B     |     -5.36|V08   |     -2.24|
|A07   |     -3.09|A16   |     -3.02|A27   |     -2.56|C     |     -4.47|V09   |     -0.75|
|A08   |     -4.69|A17   |     -0.37|A30   |      1.13|D     |     -0.76|V10   |     -2.05|

@@

For Model 3, we return to the GLM, but now we replace the categorical variables with the trained embeddings from Model
2. In other words, our model now contains only continuous variables, and each categorical factor is represented as a
scalar value and obtains its own coefficient. In contrast with the number of parameters computed in Equation
\@ref(eq:glm1), we have fewer parameters for this model:

\begin{equation}
1 + |J^{\text{num}}| + |J^{\text{cat}}|.
(\#eq:glm2)
\end{equation}

Hence, the model can be written as

\begin{equation}
\mu_i = \beta_0 + \sum_{j \in J^{\text{num}}}\beta_j x_{i,j} + \sum_{j \in J^{\text{cat}}} \beta_j z_{i,j},
\end{equation}

where $z_{i,j}$ represents the embedding value for $i$th instance of the categorical column $X_j$.

Although in this particular example we incorporate one-dimensional embeddings, one can easily include multidimensional
embeddings, which would increase the model's flexibility. The modeler would need to balance the additional lift with
some loss of interpretability, since with one-dimensional embeddings, our GLM structure dictates that there is a 
monotonic relationship between the predicted response and the input embedding, which could be desirable in certain
use cases (and undesirable in others.)

Since we only have a few parameters in this model, we exhibit the full list of fitted coefficients in Table 
\@ref(tab:glm2).

Table: (\#tab:glm2) Coefficients of Model 3.

|term                                     | estimate|
|:----------------------------------------|--------:|
|(Intercept)                              |    2.389|
|total_building_insurance_coverage        |    0.568|
|basement_enclosure_crawlspace_type       |    0.136|
|number_of_floors_in_the_insured_building |   -0.024|
|occupancy_type                           |    0.067|
|flood_zone                               |    0.035|
|primary_residence                        |   -0.001|
|community_rating_system_discount         |    0.504|

@@

Model 4 is architecturally identical to Model 2, with the only difference being the number of embedding dimensions:
instead of mapping factor levels to real numbers, we map them to points in Euclidean space, where the dimension of the 
space is the ceiling
of the factor's cardinality divided by two. Formally, in the notation of Section \ref{catdata},
$q_{\mathcal{P}^j}=\ceil{n^j_{\mathcal{P}}/2}$ for variable $X_j$. Following the practice from the previous three
models, we list the fitted embeddings for a couple variables in Tables \@ref(tab:nn2-occupancy-type) and 
\@ref(tab:nn2-flood-zone). Note that, for `flood_zone`, we are
only displaying the embedding for *one* level.

Table: (\#tab:nn2-occupancy-type) Learned embeddings for `occupancy_type` in Model 4.

|level                                       |    e1|    e2|   e3|
|:-------------------------------------------|-----:|-----:|----:|
|2 to 4 unit residential buidling            | -1.44| -1.97| 2.35|
|Non-residential building                    | -6.83| -8.11| 5.34|
|Residential building with more than 4 units | -7.63| -6.58| 5.09|
|Single family residence                     | -1.61| -2.54| 3.01|

Table: (\#tab:nn2-flood-zone) Learned embeddings for level `"A00"` of variable `flood_zone` in Model 4.

|dimension |value  |dimension |value  |dimension |value  |dimension |value  |
|:---------|:------|:---------|:------|:---------|:------|:---------|:------|
|e1        |0.020  |e11       |2.146  |e21       |-1.933 |e31       |-0.941 |
|e2        |-0.885 |e12       |1.723  |e22       |1.150  |e32       |-1.279 |
|e3        |-0.126 |e13       |0.378  |e23       |0.863  |e33       |-0.360 |
|e4        |1.491  |e14       |-0.310 |e24       |-0.471 |e34       |-1.173 |
|e5        |0.980  |e15       |2.482  |e25       |0.929  |          |       |
|e6        |0.952  |e16       |1.823  |e26       |-1.139 |          |       |
|e7        |-0.217 |e17       |-2.715 |e27       |0.113  |          |       |
|e8        |-0.611 |e18       |1.144  |e28       |1.155  |          |       |
|e9        |0.919  |e19       |-0.045 |e29       |0.707  |          |       |
|e10       |-1.196 |e20       |0.467  |e30       |1.358  |          |       |

## Visualization Techniques for Interpretation {#interpretation}

With AI and ML coming into the mainstream across industries and use cases, the topic of model explainability
has received increasing attention. In insurance pricing, @kuo2020explainability propose a framework for
model interpretability in the context of regulation and actuarial standards of practice, while @henckaerts2020boosting
investigate algorithm-specific techniques to inspect tree-based models. @kuo2020explainability and references
therein also discuss
commonly utilized model-agnostic explanation techniques, such as permutation variable importance and partial
dependence plots, that can be applied to the neural networks we consider in this paper.

In this section, we focus on ways to visualize the learned embeddings. In the authors' experiences, oftentimes
the formats of the visualizations --- basic line and scatter plots --- even more so than the content of the data, is 
helpful in explaining the concept to stakeholders.

For Model 2, the neural network with one-dimensional embeddings, the task is simple: we can simply plot the levels
on a line, as in Figure \@ref(fig:num-floors). In the case of `flood_zone`, since we happen to have the luxury of knowing a
priori the meaning of the prefixes, we can plot them separately, as in Figure \@ref(fig:flood-zone) to see how the model 
mapped them.

```{r num-floors, echo=FALSE, fig.cap="Number of floors", out.width = "80%", fig.align="center"}
knitr::include_graphics("images/num_floors.png")
```

```{r flood-zone, echo = FALSE, fig.cap="Flood zone embeddings", out.width = "60%", fig.align = "center"}
knitr::include_graphics("images/flood_zone.png")
```

For Model 4, the neural network with multi-dimensional embeddings, we would need to treat the learned values before
we can plot them. Specifically, we need to reduce the dimension of the data to 2D for embeddings with more than
two dimensions. The standard approachs for this are principal component analysis (PCA) and t-distributed stochastic
neighbor embedding (t-SNE), both of which we illustrate below.

### PCA

(TODO: insert brief theoretical stuff). See Figure \@ref(fig:flood-zone-pca).

```{r flood-zone-pca, echo = FALSE, fig.cap="Flood zone PCA", out.width = "60%", fig.align = "center"}
knitr::include_graphics("images/flood_zone_pca.png")
```

### t-SNE

(TODO: insert brief theoretical stuff). See Figure \@ref(fig:flood-zone-tsne).

```{r flood-zone-tsne, echo = FALSE, fig.cap="Flood zone t-SNE", out.width = "60%", fig.align = "center"}
knitr::include_graphics("images/flood_zone_tsne.png")
```

## Training and Inference

(TODO)

@@

# Attention based modelling {#attn}

@@

\newpage

# References {.unnumbered}

>>>>

---
title: Categorical Predictor Embeddings in Actuarial Modeling
authors:
  - name: Kevin Kuo 
    affiliation: RStudio and Kasa AI
    email: kevin@kasa.ai
  - name: Ronald Richman
    affiliation: QED Actuaries and Consultants
    email: ronaldrichman@gmail.com
abstract: |
  Enter the text of your abstract here.
keywords:
  - blah
  - blee
  - bloo
  - these are optional and can be removed
bibliography: references.bib
output:
  bookdown::pdf_book:
    base_format: rticles::arxiv_article
header-includes: |
    \usepackage{amsmath}
    \usepackage{mathtools}
    \usepackage{bm}
    \DeclarePairedDelimiter\ceil{\lceil}{\rceil}
---

@@

Recently, several studies of insurance problems have applied an alternative approach, originally from the natural langauge processing literature known as **categorical embeddings**. Instead of trying to capture the differences between categories using a single factor, the categorical embedding approach rather maps each category to a low-dimensional numeric vector, which is then used within the model as a new predictor variable.

@@

(TODO: link to github repo)

@@

```{=latex}
\begin{equation}
  z^1 = \sigma(a_1.X'+ b_1), 
\label{eq:embed_to_nn}
\end{equation}
```
where we represent the output of embedding layers concatenated together with any numerical inputs as $X'$.

@@

Attention mechanisms have been widely applied within the deep learning literature to give more flexibility to RNNs [@Bahdanau2015a] and subsequently as a replacement for FCNs in the so-called Transformer architecture, which is now widely used in natural language processing [@Vaswani2017a], and, more recently, has been applied to computer vision and tabular modelling tasks. Attention mechanisms allow deep neural nets to weight the covariates entering the model in a flexible manner; these types of applications are demonstrated in, for example, [@Vaswani2017a]. In this section we provide an introduction to attention mechanisms and then discuss applications of attention to modeling tabular data.

@@

```{=latex}
\begin{equation}
A: \mathbb{R}^{(q \times d) \times (q \times d) \times (q \times d)} \to \mathbb{R}^(q \times d), ~~
(Q, K, V) ~\mapsto ~ V^* = A(Q, K, V),
\label{eq:attention}
\end{equation}
```
where $Q \in \mathbb{R}^{q \times d}$ is a matrix of query vectors, $K \in \mathbb{R}^{q \times d}$ is a matrix of key vectors, $V \in \mathbb{R}^{q \times d}$ is a matrix of value vectors and $V^* \in \mathbb{R}^{q \times d}$, the output of the attention mechanism, is a new matrix of values. Several different options for the mapping \eqref{eq:attention} exist in the literature; here we describe the self-attention mechanism used in [@Vaswani2017a]. In more detail, consider that $V$ is a matrix of covariates relating to a regression problem, which, in the case of tabular data, will usually be composed of embeddings of the same dimension for several different categorical or numerical variables; in the case discussed here we have $q$ covariates in the form of embeddings of dimension $d$. We wish to apply the attention mechanism to $V$ to assign weights to the most important embeddings for the regression problem; this is done from the \`\`perspective" of each covariate in $V$. To do this, we formulate the query matrix $Q$ which contains information about the regression problem at hand; this is usually done by passing $V$ through a FCN trained to derive queries from the covariates. Similarly, the matrix $K$ contains the relevance of each row of $V$ for the regression problem and is also usually derived by passing each row of $V$ through another FCN. Finally, the matrix $V$ can be processed with a third FCN or left as is. The first step of the attention mechanism is to derive a matrix $A \in \mathbb{R}^{q \times q}$ of attention scores:

```{=latex}
\begin{equation}
\mathbb{R}^{(q \times d) \times (q \times d)} \to \mathbb{R}^{q \times q}, ~~
(Q, K) ~\mapsto ~ A = Q~K^T.
\label{eq:attention_scores}
\end{equation}
```
The attention scores $A$ are then normalized, or made to sum to unity, across the columns of $A$ using a soft max function i.e.

```{=latex}
\begin{equation}
A^*_{i,j} = \frac{e^{(A_{i,j})}}{\sum_{n=1}^{q} e^{(A_{i,n})}}.
\end{equation}
```
Finally, a linear combination of $V$ is formed to give the new matrix of covariates $V^*$:

```{=latex}
\begin{equation}
\mathbb{R}^{(q \times q) \times (q \times d)} \to \mathbb{R}^{q \times d}, ~~
(A^*, V) ~\mapsto ~ V^* = A^*~V.
\label{eq:attention_output}
\end{equation}
```
In most applications of attention, Equation \ref{eq:attention_scores} is modified by dividing the attention scores by $d^{\frac{1}{2}}$.

An intuitive explanation of the attention formula is as follows: we start by considering the importance of all of the covariates that we have for our modelling problem in the context of each individual covariate. This is done by assigning a score between zero and one to each covariate in the modelling problem, and we repeat this process for each of the covariates. Then, we replace each of the original covariates with a new covariate that has been formed by taking a weighted average of the covariates and the scores.

### Transformer models

Self-attention is the main building block of the current state of the art model for natural language processing, which is the Transformer model [@Vaswani2017]. The basic Transformer model applies self-attention to the covariate matrix, $X'$ which is fed into the model, to produce a new matrix of covariates, $X^*$. The two matrices are added (in an element-wise manner) together and then normalized. This matrix is then further processed through a neural network and subsequently, these processed covariates are used in the machine learning task. The Transformer model is much more accurate on NLP tasks comapred to using the raw covariates before processing with a Transformer, to the extent that this model now underlies most NLP applications that use deep learning.

### Attention Modelling for Tabular Data

Several different approaches to applying attention within the modelling of tabular data have been proposed in the literature. Perhaps the simplest way of doing this is to insert an attention layer between the embeddings and the rest of the neural network. Mathematically, we can modify Equation \ref{eq:embed_to_nn} as follows:

```{=latex}
\begin{eqnarray}
  X^* = Attn(X') \\
  z^1 = \sigma(a_1.X^* + b_1),
\label{eq:embed_to_attn_to_nn}
\end{eqnarray}
```
where $X' \in \mathbb{R}^{q \times d}$ is the matrix of embeddings after concatenating $q$ embeddings of dimension $d$ together and $X^*$ represents the matrix of embeddings $X'$ after processing with attention. Below, we demonstrate the effect of adding this simple application of attention to a tabular model.

[@Huang2020] refines this approach in a model they call the TabTransformer.

@@

These model architectures are relatively uncomplicated by today's standards, and are so chosen to better highlight the embedding components. Later, in Section X (TODO), we will investigate more involved architectures utilizing embeddings that represent the state of the art for modeling tabular data.

@@

-   Target variable: Amount paid on building claim

-   Predictors:

    -   $\log$ `building_insurance_coverage`
    -   `basement_enclosure_type`,
    -   `number_of_floors_in_the_insured_building`
    -   *prefix* of `flood_zone`
    -   `primary_residence`

-   Link function: $\log$

-   Distribution: gamma

We take the log of the continuous predictor `building_insurance_coverage`, which allows the scale of the predictor to match that of the target variable. Because the `flood_zone` variable in the original data contains 60 levels, we take the prefix of the zone code, which corresponds to the level of risk as determined by FEMA [^3]. For example, `"A01"`, `"A02"`, and so on are recoded as simply `"A"`. A log link together with the gamma distribution is a standard choice for severity modeling, which provides a multiplicative structure where the response is positive.

[^3]: <https://www.floodsmart.gov/flood-map-zone/find-yours>

@@

| variable       | label                                       | relativity |
|:---------------|:--------------------------------------------|-----------:|
| occupancy_type | Single family residence                     |       1.00 |
| occupancy_type | Non-residential building                    |       1.37 |
| occupancy_type | 2 to 4 unit residential buidling            |       1.15 |
| occupancy_type | Residential building with more than 4 units |       1.64 |
| flood_zone     | B                                           |       1.00 |
| flood_zone     | A                                           |       0.84 |
| flood_zone     | C                                           |       0.60 |
| flood_zone     | X                                           |       0.70 |
| flood_zone     | V                                           |       0.78 |
| flood_zone     | D                                           |       0.89 |

: (\#tab:relativities) Excerpt of relativities for Model 1.

@@

For Model 2, we build a feedforward neural network, whose architecture is shown in Figure \@ref(fig:nn-1). The categorical inputs go through one-dimensional embedding layers where they are each mapped to a scalar. The embeddings are then concatenated with the numeric predictors, which have been normalized in data pre-processing, before being sent to a hidden linear layer of 64 units. A ReLU (TODO: reference an equation in prev. section) activation is then applied to the outputs before they are sent to a final output layer resulting in a scalar. To constrain predictions to be positive, we apply softplus activation, which is defined to be

$$
x \mapsto \log(1 + e^x). 
$$

```{r nn-1, echo = FALSE, fig.cap = "Architecture of Models 2 and 4.", out.width = "90%", fig.align = "center"}
knitr::include_graphics("images/nn.png")
```

@@

In Tables \@ref(tab:embeddings-occupancy-type) and \@ref(tab:embeddings-flood-zone), we show the learned embeddings for the `occupancy_type` and an excerpt of them for `flood_zone`. While it may be tempting to compare these values to those in Table \@ref(tab:relativities), they represent fundamentally different concepts. It is helpful to think of the embedding values for each variable as a continuous represention of the variable itself, rather than a "coefficient" associated with the variable. Note that it is not straightforward to inspect the embeddings from a table and infer relationships to the response variable, even in a shallow network such as ours, due to nonlinearities within the model. Also note that we now have a way to incorporate the full granularity of the `flood_zone` variable, which is why more levels are shown in Table \@ref(tab:embeddings-flood-zone) as compared to Table \@ref(tab:relativities). In Section TODO, we will discuss techniques to visualize these embeddings.

| level                                       | embedding |
|:--------------------------------------------|----------:|
| 2 to 4 unit residential buidling            |      6.69 |
| Non-residential building                    |      7.16 |
| Residential building with more than 4 units |      6.33 |
| Single family residence                     |      5.43 |

: (\#tab:embeddings-occupancy-type) Learned embeddings for `occupancy_type`.

| level | embedding | level | embedding | level | embedding | level | embedding | level | embedding |
|:------|----------:|:------|----------:|:------|----------:|:------|----------:|:------|----------:|
| A     |      0.38 | A09   |     -1.71 | A18   |     -1.40 | A99   |     -2.68 | V     |     -0.41 |
| A00   |     -0.35 | A0B   |      3.85 | A19   |     -1.70 | AE    |     -4.46 | V02   |      0.65 |
| A01   |     -6.76 | A10   |     -4.03 | A20   |     -4.05 | AH    |     -2.35 | V03   |     -0.29 |
| A02   |     -4.28 | A11   |     -5.32 | A21   |     -2.08 | AHB   |     -1.71 | V04   |     -1.15 |
| A03   |     -5.93 | A12   |     -3.23 | A22   |     -2.11 | AO    |     -6.07 | V05   |     -0.50 |
| A04   |     -7.82 | A13   |      2.03 | A23   |     -1.40 | AOB   |     -4.56 | V06   |     -1.92 |
| A05   |     -3.42 | A14   |     -5.72 | A24   |      2.23 | AR    |     -0.48 | V07   |     -2.91 |
| A06   |     -4.97 | A15   |     -3.67 | A25   |      0.28 | B     |     -5.36 | V08   |     -2.24 |
| A07   |     -3.09 | A16   |     -3.02 | A27   |     -2.56 | C     |     -4.47 | V09   |     -0.75 |
| A08   |     -4.69 | A17   |     -0.37 | A30   |      1.13 | D     |     -0.76 | V10   |     -2.05 |

: (\#tab:embeddings-flood-zone) Excerpt of learned embeddings for `flood_zone`.

@@

For Model 3, we return to the GLM, but now we replace the categorical variables with the trained embeddings from Model 2. In other words, our model now contains only continuous variables, and each categorical factor is represented as a scalar value and obtains its own coefficient. (TODO: insert formal eqn for mean)

@@

Model 4 is architecturally identical to Model 2, with the only difference being the number of embedding dimensions: instead of mapping factor levels to real numbers, we map them to a point in Euclidean space, where the dimension of the space is the ceiling of the factor's cardinality divided by two. Formally, in the notation of Section \ref{catdata}, $q_{\mathcal{P}^j}=\ceil{n^j_{\mathcal{P}}/2}$ for variable $X_j$.

## Model Training and Results

To evaluate these baseline models and other models in this paper, we perform 10-fold cross validation and compare the root mean square error (RMSE) on the predicted and actual paid amounts. (somehwere we should say we're not doing a kaggle competition here and the performance metrics are really meant to show that the models aren't complete shit) (maybe a couple boxplots showing metrics for baseline models)

## Interpretations {#interpretation}

-   model explainability is important in insurance
-   the more we can learn about how we arrive at our models the better
-   cite some existing work on explainable models in actuarial literature to show people care
-   refer back to some definitions from previous sections
-   show some embedding weights from our models - 1d line plot
-   (need a multidimensional embedding for this) do some pca/tsne
-   stress that relationship of embedding weights to response variable may be nonlinear depending on architecture

@@

# Attention based modelling

@@
